A demonstration of the \LaTeXe\ class file for \itshape{SAGE Publications}

### 2015-12-01 body/efort.tex
debut planner

\section{From a guide path to a discrete sequence of contact configurations (Stage 2)}
\label{sec:contact}
As an input of this stage, we consider a truly feasible guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the root of the robot $R$. We now consider the second problem of computing a discrete sequence of contact configurations $\mathbf{Q}^{\overline{0}}$ for the limbs of the robot along $\mathbf{q}^0(t)$. From true feasibility we know that such a trajectory exists.

In this Section we first describe a single contact generation process, that is how to generate a contact configuration for a given limb 
and root location.
Then, on this method we build an iterative algorithm to generate a discrete sequence of balanced contact configurations along a truly feasible path.

\subsection{Single contact generation}
We consider a limb $R^k$  and a root configuration $\mathbf{q}^0$ truly feasible regarding $R^k$. FIGURE
From true feasibility we know that there exists an infinite list of mappings $P = [\pi_0,..,\pi_i, ...]$ that verify Eq.~\eqref{eq:pi} for $\mathbf{q}^0$.
The issue is then to evaluate locally any element of $P$ at $\mathbf{q}^0$.
This can be achieved as follows:
\begin{enumerate}
\item Generate randomly a valid limb configuration;
\item Project the end-effector onto the closest surface with inverse kinematics;
\item If a feasible solution is found, stop. Otherwise repeat from step 1.
\end{enumerate}

We partially trade the completeness property of this approach for a more efficient solution \citep{Tonneau2014}:
\begin{enumerate}
\item Generate offline $N$ valid sample limb configurations;
\item using the end-effector position expressed in $R$ coordinates as an index, store each sample in an octree stata structure;
\item Figure TODO, 2 and 3: At runtime, when contact creation is required, retrieve from the octree the list of samples $S = [\mathbf{q}^k_0, ..., \mathbf{q}^k_i, ...]$ such that $dist(\mathbf{p}^k(\mathbf{q_k}^k),O) < \epsilon$, with $\epsilon $ small ;
\item According to a user-defined heuristic $h$, sort the elements of $S$;
\item FIGURE TODO, 4 and 5: Select the first element of $S$, and project the configuration to contact with inverse kinematics;
\item If a feasible solution is found, stop. Otherwise remove the element from $S$ and go from step 5, or fail if $S$ is empty.
\end{enumerate}

This approach has two main advantages:
First, this allows us to select a number of candidates at a small cost, (a single proximity request), and to sort them according to a heuristic;
Secondly, thanks to the use of proximity requests, the number of iterations necessary for the inverse kinematics to converge is reduced.
Finally, it is easy to see that as $N$ grows to infinity, the completeness of the method is restored.
$N$ is a parameter that allows us to specify the trade-off between completeness and efficiency.
The reader is invited to refer to our previous work for an extensive discussion on the optimal value of $N$ \citep{Tonneau2014}.

\subsubsection{Efficient contact generation }

Concretely, any, we can use a random sampling approach combined with an inverse kinematics projector.

given a root configuration $\mathbf{q}^0$ and a limb $R^k$, we locally identify a countable set of mappings $P = [\pi_0,..,\pi_i] $ that verify Eq.~\eqref{eq:pi} for $\mathbf{q}^0$. and select the most relevant according to a heuristic $h$.

>>>>

\section{From a guide trajectory to a discrete sequence of contact configurations (Stage 2)}
\label{sec:efort}
As an input of this stage, we consider a truly feasible guide trajectory $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the root of the robot $R$. We now consider the second problem of computing a trajectory $\mathbf{q}^{\overline{0}}(t)$ for the limbs of the robot. Since we assume true feasibility, we know that such a trajectory exists. Contrary to previous works \cite{DBLP:conf/iser/EscandeKMG08,Bouyarmane2009}, the goal here is not to find any such trajectory but rather to select one with particular properties. Specifically, we show here how to build a contact sequence with a small number of contact variations and  good-efficiency and naturalness of the postures.

More precisely, any mapping $\pi$ introduced in \eqref{eq:pi} can be used to expand $\mathbf{q}^0(t)$ into a whole-body trajectory. We propose here a particular construction of $\pi$ leading to interesting contact sequences.

### 2015-12-01 body/intro.tex
debut planner

\subsection{Paper contribution and organization}
This paper presents a pragmatic approach to break the complexity of multi contact planning. With that objective in mind,
we believe that the separation between the guide trajectory and the contact sequence is the most promising direction \citep{DBLP:conf/iser/EscandeKMG08}.
However, this direction raises two theoretical questions that remain to be solved, or even to be properly formulated:
\begin{itemize}
\item The guide trajectory must satisfy a property guaranteeing the existence of a contact sequence to actuate it\endnote{This property is related to the controllability of the root actuated by the contact forces, but for discrete bounded properties.}. We call this property \textit{true feasibility}. This property has not been studied yet: the only way to validate a waypoint in the trajectory is to explicitly compute the contacts, which is computationally not reasonable \citep{Bouyarmane2009}.
\item There is an infinite combination of possible contact sequences for a given root trajectory. The selection of one particular contact sequence with interesting properties (minimum number of contact change, robustness, efficiency or naturalness) has been studied for cyclic cases \citep{Hauser06usingmotion}, but has not been efficiently applied to cluttered environments (\citeauthor{bouyarmane:lirmm-00777727, DBLP:conf/iser/EscandeKMG08} mostly randomly pick one contact sequence, leading to possibly very tedious contact sequences).
\end{itemize}

@@

The present paper is an extension of a conference paper to appear in the proceedings of the ISRR '15 conference TODO CITE.
The conference paper focuses on the theoritical formulation of the problem, and presents results obtained with virtual avatars.
This extension completes this work with an argumented discussion on the implementation of this theory to real world robots and problems.
In particular, we detail the robust balance criteria designed to ensure the equilibrium of the robot. Using HRP-2 as a running example, we describe how $C_{reach}$ is approximated to address
\Pa, and provide the pseudo code of the contact generation algorithm implemented to find balanced contact sequences.

>>>>

\subsection{Paper contribution and organization}
This paper presents a pragmatic approach to break the complexity of multi contact planning. With that objective in mind,
we believe that the separation between the guide trajectory and the contact sequence is the most promising direction \citep{DBLP:conf/iser/EscandeKMG08}.
However, this direction raises two theoretical questions that remain to be solved, or even to be properly formulated:
\begin{itemize}
\item The guide trajectory must satisfy a property guaranteeing the existence of a contact sequence to actuate it\endnote{This property is related to the controllability of the root actuated by the contact forces, but for discrete bounded properties.}. We call this property \textit{true feasibility}. This property has not been studied yet; the only way to validate a waypoint in the trajectory is to explicitly compute the contacts, which is computationally not reasonable \citep{Bouyarmane2009}, unless the scenario is limited to cyclic, quasi flat cases \citep{zucker2010optimization}.
\item There is an infinite combination of possible contact sequences for a given root trajectory. The selection of one particular contact sequence with interesting properties (minimum number of contact change, robustness, efficiency or naturalness) has been studied for cyclic cases \citep{Hauser06usingmotion}, but has not been efficiently applied to cluttered environments (\citeauthor{bouyarmane:lirmm-00777727, DBLP:conf/iser/EscandeKMG08} mostly randomly pick one contact sequence, leading to possibly very tedious contact sequences).
\end{itemize}

@@

The present paper is an extension of a conference paper to appear in the proceedings of the ISRR '15 conference TODO CITE.
The conference paper focuses on the theoritical formulation of the problem, and presents results obtained with virtual avatars.
This extension completes this work with an argumented discussion on the implementation of this theory to real world robots and problems.
In particular, we detail the robust balance criteria designed to ensure the equilibrium of the robot and provide details on the hpp implementation
of our contact generation algorithm. Furthermore, our experiments now concern real robots, namely HRP-2 and Hyq. TODO Appendix describes the generation process
for the robot.

### 2015-12-01 body/overview.tex
debut planner

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
\textbf{The environment} $O$ is defined as the union of the obstacles $O_i$ it contains.
The volume encompassing the trunk of the robot is denoted $W^0$ (Fig.~\ref{fig:character}-right: central cylinder). The reachable workspace of a limb $R^k$ is denoted $W^k$ (Fig.~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \mathbf{q}^k, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0$ the null displacement. We also define $W = \bigcup_{k=1}^{l}W^k$.
%
Finally, we define $W^k(\mathbf{q}^{0}), 0 < k \leq l$ as the volume $W^k$ translated and rotated by the rigid displacement~$\mathbf{q}^{0}$.

>>>>

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the trunk of the robot is denoted $W^0$ (Fig.~\ref{fig:character}-right: central cylinder). The reachable workspace of a limb $R^k$ is denoted $W^k$ (Fig.~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \mathbf{q}^k, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0$ the null displacement. We also define $W = \bigcup_{k=1}^{l}W^k$.
%
$W^k(\mathbf{q}^{0}), 0 < k \leq l$ is defined as the volume $W^k$ translated and rotated by the rigid displacement~$\mathbf{q}^{0}$.

\medskip
\textbf{The environment} $O$ is defined as the union of the obstacles $O_i$ it contains.
Finally, we define $dist(\mathbf{x}, O)$ as the minimal distance between a point $\mathbf{x}$ and the closest surface of the environment.
